Linear vertical shear instability in astrophysical disks: a numerical study

### 2014-12-14 linear_iso.tex
notation change, begun adiabatic notes

With these approximations we can re-write the governing equation as
\begin{align}\label{iso_ode1}
  &\frac{d}{dz}\left(\rho e^{-\ii k_x r
      \Omega^2/\Omega_k^2}\frac{dW}{dz}\right) +
  \sigma^2\left(\frac{1}{c_s^2}+\frac{k_x^2}{\Omega_k^2}\right) \rho W e^{-\ii k_x r
    \Omega^2/\Omega_k^2} \notag\\
  &=0.
\end{align}
Next, we multiply Eq. \ref{iso_ode1} by $W^{*}\exp{\left(\ii k_x r
    \Omega^2/\Omega_k\right)}$ and integrate 
vertically from $z=-\zmax$ to $z=\zmax$. We assume either $W$ or
$dW/dz$ vanishes at $z=\pm \zmax$, or that $\zmax$ is sufficiently
large so that the boundary terms are negligible because of the
decaying background density with increasing height. Then,
\begin{align}\label{neccessary_cond}
   &\sigma^2\left(\frac{1}{c_s^2}+\frac{k_x^2}{\Omega_k^2}\right)\int_{-\zmax}^{\zmax}\rho|W|^2dz \notag\\
  &=\int_{-\zmax}^{\zmax}\rho\left|\frac{dW}{dz}\right|^2dz 
  +\frac{\ii k_x r}{\Omega_k^2}\int_{-\zmax}^{\zmax}\rho\frac{d\Omega^2}{dz}W^*\frac{dW}{dz}dz. 
\end{align}
It follows that for instability ($\imag\sigma>0$), it is neccessary to
have $d\Omega^2/dz\neq 0$, i.e. vertical shear.

\subsection{Maximum growth rate}   
Here we show that the growth rate is limited by the maximum vertical
shear in the domain. It is convenient to introduce the dimensionless
vertical co-ordinate $\hat{z} = z/\Hiso$ and wavenumber
$\hat{k}=k_x\Hiso$. Then 
the integral relation Eq. \ref{neccessary_cond} becomes 
\begin{align}\label{integral_relation1}
  &\left(\frac{\sigma}{\Omega_k}\right)^2\left(1+\hat{k}^2\right)
  \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho|W|^2d\hat{z}  
  \notag\\ 
  &=\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left|\frac{dW}{d\hat{z}}\right|^2d\hat{z}      
  +\ii
  \epsilon\hat{k}q\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
  f(\hat{z}) W^*\frac{dW}{d\hat{z}}d\hat{z},    
\end{align}   
where $f(\hat{z})$ is defined such that
\begin{align}\label{fz_shear}
  \frac{d\Omega^2}{d\hat{z}} = \epsilon^2q f(\hat{z})\Omega_k^2.
\end{align}

For convenience, we write $\hat{\sigma}\equiv\sigma/\Omega_k = \hat{\omega} +
\ii\hat{\nu}$ with $\hat{\omega},\hat{\nu}$ being the dimensionless
mode frequency and growth rates. Then the real and imaginary parts of
Eq. \ref{integral_relation1} are
\begin{align}
  &\left(\hat{\omega}^2-\hat{\nu}^2\right)\left(1+\hat{k}^2\right)
   \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho|W|^2d\hat{z} -
  \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left|\frac{dW}{d\hat{z}}\right|^2d\hat{z}
  \notag\\
  &=\real\left[\ii
  \epsilon\hat{k}q\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
 f(\hat{z}) W^*\frac{dW}{d\hat{z}}d\hat{z}\right], \notag\\
& 2\hat{\omega}\hat{\nu}\left(1+\hat{k}^2\right)
   \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho|W|^2d\hat{z}\notag\\
   &=\imag\left[\ii
  \epsilon\hat{k}q\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
 f(\hat{z}) W^*\frac{dW}{d\hat{z}}d\hat{z}\right].
\end{align}
Adding the square of these equations give
\begin{align}
&\left[|\hat{\sigma}|^2\left(1+\hat{k}^2\right)
   \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho|W|^2d\hat{z} -
  \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left|\frac{dW}{d\hat{z}}\right|^2d\hat{z}\right]^2\notag\\
&+4\hat{\nu}^2\left(1+\hat{k}^2\right) 
  \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
   |W|^2d\hat{z}\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left|\frac{dW}{d\hat{z}}\right|^2d\hat{z}\notag\\
   &=\left|\ii
  \epsilon\hat{k}q\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
 f(\hat{z}) W^*\frac{dW}{d\hat{z}}d\hat{z}\right|^2.
\end{align}
It is clear that
\begin{align}\label{sigma_finite_domain} 
&4\hat{\nu}^2\left(1+\hat{k}^2\right) 
  \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
   |W|^2d\hat{z}\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left|\frac{dW}{d\hat{z}}\right|^2d\hat{z}\notag\\
   &\leq\left|
  \epsilon\hat{k}q\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
 f(\hat{z}) W^*\frac{dW}{d\hat{z}}d\hat{z}\right|^2.
\end{align}
% We now take the same approach as in the previous section. We first
% assume an neutral solution corresponding to $q\equiv0$, for which
% $\sigma$ is real and we can take $W$ to be real as well. We then
% set $q\to\delta q$ and linearize. This procedure means
% \begin{align}
%   &\left(\frac{\sigma}{\Omega_k}\right)^2\left(1+\hat{k}^2\right)
%   \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho W^2d\hat{z}  
%   =\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left(\frac{dW}{d\hat{z}}\right)^2d\hat{z},\\ 
%   &\frac{2\sigma\imag{\delta\sigma}}{\Omega_k^2}\left(1+\hat{k}^2\right)
%   \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho W^2d\hat{z}\notag\\
%   &=\epsilon\hat{k}\delta q\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
%   f(\hat{z}) W\frac{dW}{d\hat{z}}d\hat{z}. 
% \end{align}
% Eliminating $\sigma^2$ between the above equations, we obtain
% \begin{align}\label{sigma_finite_domain}
%   &4
%   \left[\frac{\imag{\left(\delta\sigma\right)}}{\Omega_k}\right]^2\left(1+\hat{k}^2\right) 
%   \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
%   W^2d\hat{z}\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left(\frac{dW}{d\hat{z}}\right)^2d\hat{z}\notag\\
%   &=\left(\epsilon\hat{k}\delta q\right)^2\left(\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
%   f(\hat{z}) W\frac{dW}{d\hat{z}}d\hat{z}\right)^2. 
% \end{align}
On the left hand side of this inequality, we apply the Cauchy-Schwarz
inequality to obtain
\begin{align}
  &\left( \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
    |W|\left|\frac{dW}{d\hat{z}}\right|d\hat{z}\right)^2\notag\\&\leq
  \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho 
  |W|^2d\hat{z}\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left|\frac{dW}{d\hat{z}}\right|^2d\hat{z}.
\end{align}
On the right hand side of Eq. \ref{sigma_finite_domain} we have
\begin{align}
  \left|\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
    f(\hat{z}) W^*\frac{dW}{d\hat{z}}d\hat{z}\right|\leq \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
  \left|f(\hat{z})W^*\frac{dW}{d\hat{z}}\right|d\hat{z} \notag\\
  \leq
  \mathrm{max}\left(|f|\right)\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
  |W|\left|\frac{dW}{d\hat{z}}\right|d\hat{z},
\end{align}
where $\mathrm{max}(|f|)$ is the maximum value of $|f|$ in
$z\in[-\zmax,\zmax]$. Inserting these inequalities into
Eq. \ref{sigma_finite_domain} gives
\begin{align}\label{max_growth}
  \hat{\nu}\equiv\left|\frac{\imag{(\sigma)}}{\Omega_k}\right| \leq
  \frac{\epsilon |\hat{k} q|}{2\sqrt{1+\hat{k}^2}}\mathrm{max}(|f|). 
\end{align}
Recalling that $f$ represents vertical shear (Eq. \ref{fz_shear}), it
follows that the maximum possible growth rate of unstable modes,
satisfying the above boundary conditions, is limited by the maximum
value of the vertical shear in the domain considered.

@@

\subsection{Thin-disk limit}
We can make further progress for thin-disks ($\epsilon\ll1$), 
so we may expand the background density and vertical shear in powers
of $z/r$. To lowest order we obtain 
\begin{align}
  &\rho(\hat{z}) \simeq \rho_{00}
  \exp{\left(-\frac{\hat{z}^2}{2}\right)} \equiv \rho_{00}w(\hat{z}),\label{thin_dens}\\
  &\frac{d\Omega^2}{d\hat{z}} \simeq \epsilon^2q\Omega_k^2\hat{z}. \label{thin_vshear}
\end{align}
With this approximation and those in \S\ref{integral_relation}, the
governing equation becomes 
\begin{align}\label{iso_ode3}
  \frac{d^2W}{d\hat{z}^2} - \left(1 + \ii q\epsilon
    \hat{k}\right)\hat{z}\frac{dW}{d\hat{z}} +
  \left(\frac{\sigma}{\Omega_k}\right)^2\left(1+\hat{k}^2\right)W = 
  0.
\end{align}
We remark that for large $|\hat{k}|$, Eq. \ref{iso_ode3} is the same as
that derived by \cite{nelson13}, although we have taken a different
route.

To complete the problem we must specify boundary conditions. The
maximum vertical domain size we should consider is limited by the thin disk
approximation (i.e. $z\lesssim r$). It is most common to take $\zmax\to\infty$
and impose that the kinetic  energy remain bounded at infinity.  This
is appropriate for the density because both the true density profile
and its thin-disk approximation decay rapidly away from the midplane.

However, the thin-disk representation of vertical shear diverges with 
$\hat{z}$, unlike the true vertical shear profile which eventually
decays.  Thus taking $\zmax\to$ will permit unphysically large growth
rates (as demonstrated below). Nevertheless, the infinite disk is
useful to consider since it permits an analytical discussion.

\subsection{Stability in the absence of vertical shear}
When $q=0$, Eq. \ref{iso_ode3} is
\begin{align}\label{hermite_ode}
  \frac{d}{d\hat{z}}\left[w(\hat{z})\frac{dW}{d\hat{z}}\right] + nW
  w(\hat{z}) =0, 
\end{align}
where we have defined a new eigenvalue
\begin{align}
  n \equiv \left(\frac{\sigma}{\Omega_k}\right)^2(1+\hat{k}^2). 
\end{align} 
Eq. \ref{iso_ode3} is Hermite's differential equation. If we impose
that the kinetic energy density remain bounded at infinity, then  
\begin{align}
  W \propto \He_n(\hat{z}),
\end{align}
and $n$ is a non-negative integer. This translates to a real
eigenfrequency $\sigma$ such that
\begin{align}
  \left|\frac{\sigma}{\Omega_k}\right| = \sqrt{n}
  \left(1+\hat{k}^2\right)^{-1/2}. 
\end{align}
Since we have assumed $|\sigma^2|\ll \kappa^2$, our analysis is only
valid for large wavenumbers such that $\hat{k}^2\gg 
n-1$. Physically, this corresponds to radial length-scales much
smaller than the local disk scale height.

\subsection{Polynomial solutions with vertical shear}
We can seek power-series solution to Eq. \ref{iso_ode3},
\begin{align}
  W(\zhat) = \sum_{m=0}^\infty a_m\zhat^m. 
\end{align}
Then the coefficients must satisfy the recurrence relation
\begin{align}
  (m+2)(m+1)a_{m+2} +
  \left[n - m\left(1+\ii q h \hat{k}\right)\right] a_m = 0, 
\end{align}
where $n$ is not neccessarily an integer. Indeed, if we demand
a polynomial of degree $M$ as the solution, then the eigenfrequency
must satisfy
\begin{align}
\left(\frac{\sigma}{\Omega_k}\right)^2 = M\left(\frac{1+\ii q \epsilon
    \hat{k}}{1+\hat{k}^2}\right).
\end{align}
The eigenfrequency $\sigma$ is therefore complex. We are interested in
unstable modes for which $\imag{\sigma}$ is the growth rate and is
given by 
\begin{align}\label{simple_growth}
   \frac{\imag(\sigma)}{\Omega_k} =\sqrt{
   \frac{M}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2\epsilon^2\hat{k}^2} - 
    1\right)}. 
\end{align}
For fixed $q$ and $\epsilon$, the growth rate vanishes for both
$\hat{k}^2\to0$ and $\hat{k}^2\to\infty$. The maximum growth rate
occurs at the optimum wavenumber $\hat{k}_\mathrm{opt}$
\begin{align}
  |\hat{k}_\mathrm{opt}| = \sqrt{\frac{2+|\epsilon q|}{|\epsilon q|}},
\end{align}
and
\begin{align}
\mathrm{max}\left[\imag(\sigma)\right] =\frac{\sqrt{M}|\epsilon
  q|}{2\sqrt{1+|\epsilon q|}}\Omega_k. 
\end{align}

@@

\subsection{Instability in the presence of weak vertical shear}
We can also investigate the destabilizing effect of vertical shear
without explicitly solving the full governing equation,
Eq. \ref{iso_ode3}. We first
consider a system with $q\equiv0$, for which the eigenfunctions are
Hermite polynomials and the real eigenfrequencies are known. We then
perturb this system by introducing a weak vertical shear ($|q|\ll1$)
and linearize the governing equation. This procedure is
\begin{align}   
  &q \to 0 + \delta q\\
  &W \to \He_n + \delta V \\
  &\sigma \to \sigma + \delta\sigma, 
\end{align}
where for this exercise we restore $\sigma$ as the eigenfrequency.

Linearizing the integral relation Eq. \ref{integral_relation1} in the
thin-disk limit, and taking the
imaginary part, we have
\begin{align}
  \frac{2\sigma\imag(\delta\sigma)}{\Omega_k^2}
  \left(1+\hat{k}^2\right) \int_{-\infty}^{\infty} w(\zhat)
  \He_n^2(\hat{z}) d\hat{z} \notag\\
  = \delta q \epsilon \hat{k} 
  \int_{-\infty}^{\infty}
  w(\zhat)\hat{z}\He_n(\hat{z})\frac{d\He_n}{d\zhat} d\hat{z}
\end{align}
Recognizing $\hat{z} = \He_1$ and using $d\He_n/d\hat{z} = n
\He_{n-1}$ we have
 \begin{align}
   \frac{2\sigma\imag(\delta\sigma)}{\Omega_k^2}
   \left(1+\hat{k}^2\right) \int_{-\infty}^{\infty} w(\zhat)
   \He_n^2(\hat{z}) d\hat{z} \notag\\
   =\delta q \epsilon \hat{k} 
   \int_{-\infty}^{\infty}
   w(\zhat)\He_1(\hat{z})\He_n(\hat{z})n\He_{n-1}(\hat{z})d\hat{z}. 
 \end{align}

@@

 \begin{align}
  \frac{2\sigma\imag(\delta\sigma)}{\Omega_k^2}
  \left(1+\hat{k}^2\right) = \delta q \epsilon \hat{k} n.
 \end{align}
Inserting the original eigenfrequency $\sigma = \pm
\sqrt{n}\Omega_k(1+\hat{k}^2)^{-1/2}$ gives
\begin{align}
  \frac{\imag(\delta\sigma)}{\Omega_k} = \pm \frac{1}{2}\delta q \epsilon
  \sqrt{n} \frac{\hat{k}}{\sqrt{1+\hat{k}^2}}. 
\end{align}
This result agrees with Eq. \ref{simple_growth} in the limit
$|q|\to0$. 
%for $n=1$, since the
%function $\He_1$ solves the governing equation exactly. 
For $\hat{k}\gg 1$ we have
\begin{align}
  \frac{\imag(\delta\sigma)}{\Omega_k} \simeq \pm \frac{1}{2}\delta q \epsilon
  \sqrt{n} \sgn{\hat{k}}. 
\end{align}

>>>>

With these approximations and introducing the dimensionless variables
\begin{align}
  \hat{z} = z/\Hiso,\quad \hat{k}=k_x\Hiso, \quad \hat{\sigma} = \sigma/\Omega_k,
\end{align}
the governing equation can be re-written as
\begin{align}
  0 = W^{\prime\prime} + \left[\ln\rho^\prime - \ii \epsilon q \hat{k}
    f(\zhat)\right]W^\prime + \hat{\sigma}^2\left(1+\hat{k}^2\right)W,\label{iso_ode1}
\end{align}
where the primes denote $d/d\zhat$ and $f(\zhat)$ is defined such that
\begin{align}\label{fz_shear}
  \frac{d\Omega^2}{d\hat{z}} = \epsilon^2q f(\hat{z})\Omega_k^2.
\end{align}

Next, we multiply Eq. \ref{iso_ode1} by $\rho W^*$ and integrate 
vertically from $\zhat=\zhat_1$ to $z=\zhat_2$. We neglect boundary
terms when integrating by parts, by assuming $W$ or
$W^\prime$ vanishes at the boundaries, or that the boundaries are 
sufficiently far away so that the boundary terms are negligible because of the
decaying background density with increasing height. Then,
\begin{align}
  &\hat{\sigma}^2\left(1+\hat{k}^2\right)\int_{\zhat_1}^{\zhat_2}\rho|W|^2d\zhat \notag\\
  &=\int_{\zhat_1}^{\zhat_2}\rho|W^\prime|^2d\zhat 
  +\ii \epsilon q \hat{k}\int_{\zhat_1}^{\zhat_2}\rho f(\zhat) W^*W^\prime d\zhat.\label{integral_relation1}
\end{align}
It follows that for instability ($\imag\hat{\sigma}>0$), it is neccessary to
have $q\neq0$ or more generally $d\Omega^2/dz\neq 0$, i.e. vertical shear.

% With these approximations we can re-write the governing equation as
% \begin{align}\label{iso_ode1}
%   &\frac{d}{dz}\left(\rho e^{-\ii k_x r
%       \Omega^2/\Omega_k^2}\frac{dW}{dz}\right) +
%   \sigma^2\left(\frac{1}{c_s^2}+\frac{k_x^2}{\Omega_k^2}\right) \rho W e^{-\ii k_x r
%     \Omega^2/\Omega_k^2} \notag\\
%   &=0.
% \end{align}
% Next, we multiply Eq. \ref{iso_ode1} by $W^{*}\exp{\left(\ii k_x r
%     \Omega^2/\Omega_k\right)}$ and integrate 
% vertically from $z=-\zmax$ to $z=\zmax$. We assume either $W$ or
% $dW/dz$ vanishes at $z=\pm \zmax$, or that $\zmax$ is sufficiently
% large so that the boundary terms are negligible because of the
% decaying background density with increasing height. Then,
% \begin{align}\label{neccessary_cond}
%    &\sigma^2\left(\frac{1}{c_s^2}+\frac{k_x^2}{\Omega_k^2}\right)\int_{-\zmax}^{\zmax}\rho|W|^2dz \notag\\
%   &=\int_{-\zmax}^{\zmax}\rho\left|\frac{dW}{dz}\right|^2dz 
%   +\frac{\ii k_x r}{\Omega_k^2}\int_{-\zmax}^{\zmax}\rho\frac{d\Omega^2}{dz}W^*\frac{dW}{dz}dz. 
% \end{align}
% It follows that for instability ($\imag\sigma>0$), it is neccessary to
% have $d\Omega^2/dz\neq 0$, i.e. vertical shear.

\subsection{Maximum growth rate}   
Here we show that the growth rate is limited by the maximum vertical
shear in the domain. For convenience, we write $\hat{\sigma} = \hat{\omega} +
\ii\hat{\nu}$ with $\hat{\omega},\hat{\nu}$ being the real dimensionless
mode frequency and growth rates. Then the real and imaginary parts of
Eq. \ref{integral_relation1} are
\begin{align}
  &\left(\hat{\omega}^2-\hat{\nu}^2\right)\left(1+\hat{k}^2\right)
  \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z} -
  \int_{\zhat_1}^{\zhat_2}\rho|W^\prime|^2d\hat{z}
  \notag\\
  &=\real\left[\ii
    \epsilon\hat{k}q \int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\zhat\right], \notag\\
& 2\hat{\omega}\hat{\nu}\left(1+\hat{k}^2\right)
   \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z}\notag\\
   &=\imag\left[\ii
     \epsilon\hat{k}q \int_{\zhat_1}^{\zhat_2}\rho
     f(\hat{z}) W^*W^\prime d\hat{z}\right].
\end{align}
Adding the square of these equations give
\begin{align}
&\left[|\hat{\sigma}|^2\left(1+\hat{k}^2\right)
   \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z} -
  \int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}\right]^2\notag\\
&+4\hat{\nu}^2\left(1+\hat{k}^2\right) 
  \int_{\zhat_1}^{\zhat_2}\rho
   |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}\notag\\
   &=\left|\ii
  \epsilon\hat{k}q\int_{\zhat_1}^{\zhat_2}\rho
 f(\hat{z}) W^*W^\prime d\hat{z}\right|^2.
\end{align}
It is clear that
\begin{align}\label{sigma_finite_domain} 
&4\hat{\nu}^2\left(1+\hat{k}^2\right) 
  \int_{\zhat_1}^{\zhat_2}\rho
   |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}\notag\\
   &\leq\left|
  \epsilon\hat{k}q\int_{\zhat_1}^{\zhat_2}\rho
 f(\hat{z}) W^*W^\prime d\hat{z}\right|^2.
\end{align}
On the left hand side of this inequality, we apply the Cauchy-Schwarz
inequality to obtain
\begin{align}
  &\left( \int_{\zhat_1}^{\zhat_2}\rho
    |W|\left|W^\prime \right|d\hat{z}\right)^2\leq
  \int_{\zhat_1}^{\zhat_2}\rho 
  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}.
\end{align}
On the right hand side of Eq. \ref{sigma_finite_domain} we have
\begin{align}
  \left|\int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\hat{z}\right|\leq \int_{\zhat_1}^{\zhat_2}\rho
  \left|f(\hat{z})W^*W^\prime \right|d\hat{z} \notag\\
  \leq
  \mathrm{max}\left(|f|\right)\int_{\zhat_1}^{\zhat_2}\rho
  |W|\left|W^\prime \right|d\hat{z},
\end{align}
where $\mathrm{max}(|f|)$ is the maximum value of $|f|$ in
$\zhat\in[\zhat_1,\zhat_2]$. Inserting these inequalities into
Eq. \ref{sigma_finite_domain} gives
\begin{align}\label{max_growth}
  |\hat{\nu}|\leq
  \frac{\epsilon |\hat{k} q|}{2\sqrt{1+\hat{k}^2}}\mathrm{max}(|f|). 
\end{align}
Recalling that $f$ represents vertical shear (Eq. \ref{fz_shear}), it
follows that the maximum possible growth rate of unstable modes,
satisfying the above boundary conditions, is limited by the maximum
value of the vertical shear in the domain considered.

@@

\subsection{Thin-disk limit}
We can make further progress for thin-disks ($\epsilon\ll1$), 
so we may expand the background density and vertical shear in powers
of $z/r$. To lowest order we obtain 
\begin{align}
  &\rho(\hat{z}) \simeq \rho_{00}
  \exp{\left(-\frac{\hat{z}^2}{2}\right)} \equiv \rho_{00}w(\hat{z}),\label{thin_dens}\\
  &\frac{d\Omega^2}{d\hat{z}} \simeq \epsilon^2q\Omega_k^2\hat{z}. \label{thin_vshear}
\end{align}
With this approximation and those in \S\ref{integral_relation}, the
governing equation becomes 
\begin{align}\label{iso_ode3}
  W^{\prime\prime} - \left(1 + \ii q\epsilon
    \hat{k}\right)\hat{z}W^\prime  +
  \hat{\sigma}^2\left(1+\hat{k}^2\right)W = 
  0.
\end{align}
We remark that for large $|\hat{k}|$, Eq. \ref{iso_ode3} is the same as
that derived by \cite{nelson13}, although we have taken a different
route.

To complete the problem we must specify boundary conditions. The
maximum vertical domain size we should consider is limited by the thin disk
approximation (i.e. $z\lesssim r$). It is common to take $\mathrm{max}|z|\to\infty$
and impose that the kinetic  energy remain bounded at infinity.  This
is appropriate for the density because both the true density profile
and its thin-disk approximation decay rapidly away from the midplane.

However, the thin-disk representation of vertical shear diverges with 
$\hat{z}$, unlike the true vertical shear profile which eventually
decays.  Thus, taking $\zhat_1\to -\infty$ and $\zhat_2\to\infty$ will permit unphysically large growth
rates (as demonstrated below). Nevertheless, the infinite disk is
useful to consider since it permits an analytical discussion.

\subsection{Stability in the absence of vertical shear}
When $q=0$, Eq. \ref{iso_ode3} is
\begin{align}\label{hermite_ode}
  \frac{d}{d\hat{z}}\left[w(\hat{z})W^\prime \right] + nW
  w(\hat{z}) =0, 
\end{align}
where we have defined a new eigenvalue
\begin{align}
  n \equiv \hat{\sigma}^2(1+\hat{k}^2). 
\end{align} 
Eq. \ref{iso_ode3} is Hermite's differential equation. If we impose
that the kinetic energy density remain bounded at infinity, then  
\begin{align}
  W \propto \He_n(\hat{z}),
\end{align}
and $n$ is a non-negative integer. This translates to a real
eigenfrequency $\sigma$ such that
\begin{align}
  \left|\hat{\sigma}\right| = \sqrt{n}
  \left(1+\hat{k}^2\right)^{-1/2}. 
\end{align}
Since we have assumed $|\sigma^2|\ll \kappa^2$, our analysis is only
valid for large wavenumbers such that $\hat{k}^2\gg 
n-1$. Physically, this corresponds to radial length-scales much
smaller than the local disk scale height.

\subsection{Polynomial solutions with vertical shear}
We can seek power-series solution to Eq. \ref{iso_ode3},
\begin{align}
  W(\zhat) = \sum_{m=0}^\infty a_m\zhat^m. 
\end{align}
Then the coefficients must satisfy the recurrence relation
\begin{align}
  (m+2)(m+1)a_{m+2} +
  \left[n - m\left(1+\ii \epsilon q  \hat{k}\right)\right] a_m = 0, 
\end{align}
where $n$ is not neccessarily an integer. Indeed, if we demand
a polynomial of degree $M$ as the solution, then the eigenfrequency
must satisfy
\begin{align}
\hat{\sigma}^2 = M\left(\frac{1+\ii q \epsilon
    \hat{k}}{1+\hat{k}^2}\right).
\end{align}
The eigenfrequency $\sigma$ is therefore complex. We are interested in
unstable modes for which $\imag{\hat{\sigma}}>0$ is the growth rate and is
given by 
\begin{align}\label{simple_growth}
  \hat{\nu} =\sqrt{
   \frac{M}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2\epsilon^2\hat{k}^2} - 
    1\right)}. 
\end{align}
For fixed $q$ and $\epsilon$, the growth rate vanishes for both
$\hat{k}^2\to0$ and $\hat{k}^2\to\infty$. The maximum growth rate
occurs at the optimum wavenumber $\hat{k}_\mathrm{opt}$
\begin{align}
  |\hat{k}_\mathrm{opt}| = \sqrt{\frac{2+|\epsilon q|}{|\epsilon q|}},
\end{align}
and
\begin{align}
\mathrm{max}\left(\hat{\nu}\right) =\frac{\sqrt{M}|\epsilon
  q|}{2\sqrt{1+|\epsilon q|}}. 
\end{align}

@@

\subsection{Instability in the presence of weak vertical shear}
We can also investigate the destabilizing effect of vertical shear
without explicitly solving the full governing equation,
Eq. \ref{iso_ode3}. We first
consider a system with $q\equiv0$, for which the eigenfunctions are
Hermite polynomials and the real eigenfrequencies are known. We then
perturb this system by introducing a weak vertical shear ($|q|\ll1$)
and linearize the governing equation. This procedure is
\begin{align}   
  &q \to 0 + \delta q\\
  &W \to \He_n + \delta W \\
  &\hat{\sigma} \to \hat{\sigma} + \delta\hat{\sigma}. 
\end{align}
%where for this exercise we restore $\sigma$ as the eigenfrequency.

Linearizing the integral relation Eq. \ref{integral_relation1} in the
thin-disk limit, and taking the
imaginary part, we have
\begin{align}
  2\hat{\sigma}\imag(\delta\hat{\sigma})
  \left(1+\hat{k}^2\right) \int_{-\infty}^{\infty} w(\zhat)
  \He_n^2(\hat{z}) d\hat{z} \notag\\
  = \delta q \epsilon \hat{k} 
  \int_{-\infty}^{\infty}
  w(\zhat)\hat{z}\He_n(\hat{z})\He_n^\prime(\zhat) d\hat{z}
\end{align}
Recognizing $\hat{z} = \He_1(\hat{z})$ and using $\He_n^\prime = n
\He_{n-1}$ we have
 \begin{align}
   2\hat{\sigma}\imag(\delta\hat{\sigma})
   \left(1+\hat{k}^2\right) \int_{-\infty}^{\infty} w(\zhat)
   \He_n^2(\hat{z}) d\hat{z} \notag\\
   =\delta q \epsilon \hat{k} 
   \int_{-\infty}^{\infty}
   w(\zhat)\He_1(\hat{z})\He_n(\hat{z})n\He_{n-1}(\hat{z})d\hat{z}. 
 \end{align}

@@

 \begin{align}
   2\hat{\sigma}\imag(\delta\hat{\sigma})
  \left(1+\hat{k}^2\right) = \delta q \epsilon \hat{k} n.
 \end{align}
Inserting the original eigenfrequency $\hat{\sigma} = \pm
\sqrt{n}(1+\hat{k}^2)^{-1/2}$ gives
\begin{align}
  \imag(\delta\hat{\sigma})= \pm \frac{1}{2}\delta q \epsilon
  \sqrt{n} \frac{\hat{k}}{\sqrt{1+\hat{k}^2}}. 
\end{align}
This result agrees with Eq. \ref{simple_growth} in the limit
$|q|\to0$. 
%for $n=1$, since the
%function $\He_1$ solves the governing equation exactly. 
For $\hat{k}\gg 1$ we have
\begin{align}
  \imag(\delta\hat{\sigma})\simeq \pm \frac{1}{2}\delta q \epsilon
  \sqrt{n} \sgn{\hat{k}}. 
\end{align}
